---
title: "Column Export 2015-04-04"
author: "Toph Allen"
date: "April 4, 2015"
output: html_document
---

## Introduction

I'm extracting data from the DFHC Excel sheet for the following variables:

1. Proportion (+- sample size for question) of respondents indicating that they have:
    - butchered live animals in the last year
    - seen wildlife near their home in the last year & what types
    - seen wildlife inside their home in the last year
    - seen rodents in their home in the last year
    - been bitten or scratched by wildlife in the last year
    - entered the forest and why
2. Types of wildlife consumed (presence/absence matrix by respondent ID)

## Setup

```{r knitr_options}
knitr::opts_knit$set(root.dir = "~/Dropbox (EHA)/repositories/dfhc", width = 75)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE)
options(digits = 2)

```

```{r r_setup}
library(stringr)
library(plyr)
library(dplyr)
library(magrittr)
library(printr)

source("R/helper_functions.R")

descending_table <- function(x) {
  dplyr::arrange(data.frame(table(x)), -Freq)
}

descending_kable <- function(x) {
  knitr::kable(descending_table(x))
}
```

## Loading raw dataset, selecting columns

Looking over the list of things Lizzie asked me to extract and comparing it to the spreadsheet, I came up with the following columns:

- metadata
    - 6: respondent gender
- wildlife contact
    - 22: wildlife near home, yes/no
    - 24: wildlife near home, taxonomy categories
    - 26: wildlife in home, yes/no
    - 28: wildlife in home, taxonomy categories
    - 30: wildlife in home, frequency categories
    - 132: wildlife contact, categories
    - 134: where contact happened, categories
    - 136: wildlife species, taxonomy categories
- forest entry
    - 78: enter forest - why: self, categories
    - 76: enter forest - how often, frequency categories
- wildlife consumption
    - 122: self past consumption wildlife, yes/no
    - 123: people in community past consumption wildlife, yes/no
    - 124: self:wildlife currently consumed, taxonomy categories
    - 126: community: wildlife currently consumed, taxonomy categories
- butchering live animals in the past year
    - N.B. “Butchered live animals in the last year” isn’t a yes/no question. It seems to be spread across multiple categories, and the answers don’t seem to indicate whether the person or someone else butchered the animal.
    - “Who prepares the carcass” only has two answers in the entire column.

I'm going to not bother with the last set of questions until I can check with Lizzie and Allison about how to get that information.

```{r load_and_select_data}

dfhc_raw <- read.csv("rawdata/predict_hac_data_raw.csv",
                 as.is = TRUE)

dfhc_raw <- dfhc_raw[1:406, ]

cols_to_keep <- c("pid" = 1,
                  "respondent_gender" = 6,
                  "wildlife_near_home_yn" = 22,
                  "wildlife_near_home_tax_cat" = 24,
                  "wildlife_in_home_yn" = 26,
                  "wildlife_in_home_tax_cat" = 28,
                  "wildlife_in_home_freq_cat" = 30,
                  "wildlife_contact_cat" = 132,
                  "where_contact_happened_cat" = 134,
                  "wildlife_species_tax_cat" = 136,
                  "enter_forest_how_often_freq_cat" = 76,
                  "enter_forest_why_self_cat" = 78,
                  "wildlife_consumpt_self_past_yn" = 122,
                  "wildlife_consumpt_community_past_yn" = 123,
                  "wildlife_consumpt_self_curr_cat" = 124,
                  "wildlife_consumpt_self_curr_free" = 125,
                  "wildlife_consumpt_community_curr_cat" = 126,
                  "wildlife_consumpt_community_curr_free" = 127)

cols_raw <- dfhc_raw[, cols_to_keep]

names(cols_raw) <- names(cols_to_keep)

cols_raw <- colwise(initial_character_column_cleaning)(cols_raw)
```

How many distinct responses each question has will inform how much cleaning we'll need to do. We'll also print out tables of the responses.

```{r assess_cleaning}

str(dplyr::summarise_each(cols_raw, funs(n_distinct)))

for (name in names(cols_raw[-1])) {
  cat("\n\n")
  print(name)
  print(descending_kable(cols_raw[, name]))
}

```

For each column:

- Male/female seems good.
- Wildlife near home yes/no seems good.
- Wildlife near home categories needs tokenization and turning into a matrix.
- Wildlife in home yes/no seems good.
- Wildlife in home categories needs tokenization and turning into a matrix.
- Wildlife in home frequency needs… well, it looks like there are one or two answers where "daily" appears alongside another category. Those just need to be replaced with "daily".
- Wildlife contact category has a few entries where "bitten" occurs alongside other entries. I'm just going to add those to "bitten".
- Wildlife contact where happened has only two answers which aren't in the proper categories, and both could be replaced with "in the home".
- Wildlife contact species taxonomic categories needs splitting just a bit of recategorization and renaming.
- Wildlife contact category has a few responses where "few times per week" etc. appear alongside "other" or "never". These could be recategorized. But also, they're only one or two responses.
- Why enter forest reasons are a little confusing, and I have questions about the best way to lump it into bigger categories.

- Wildlife consumed (past) yes/no are good.

- The "wildlife currently consumed" questions aren't yes/no, they're categories, which I imagine could change the way that people answer and will make them non-comparable to the other questions.
- Taxonomy of wildlife consumed is basically useless; there are only seven non-blank responses.
- Same for the community-related question. There are eight responses.
- However, the "other" columns here could be fruitful, as it seems that almost every "other" answer is a "yes", specifying animals. They will need to be tokenized before analysis.

So, this is the data cleaning that will happen:

- Export as-is:
    - `pid`
    - `respondent_gender`
    - `wildlife_near_home_yn`
    - `wildlife_in_home_yn`
    - `wildlife_consumpt_self_past_yn`
    - `wildlife_consumpt_community_past_yn`

## Making a pretty list of data cleaning to apply to each column.

```{r}
cols_meta <- list("pid" = "as_is",
                  "respondent_gender" = "as_is",
                  "wildlife_near_home_yn" = "as_is",
                  "wildlife_near_home_tax_cat" = c("tokenize", "replace"),
                  "wildlife_in_home_yn" = "as_is",
                  "wildlife_in_home_tax_cat" = c("tokenize", "replace"),
                  "wildlife_in_home_freq_cat" = "replace",
                  "wildlife_contact_cat" = "replace",
                  "where_contact_happened_cat" = "replace",
                  "wildlife_species_tax_cat" = "replace",
                  "enter_forest_how_often_freq_cat" = "replace",
                  "enter_forest_why_self_cat" = "replace",
                  "wildlife_consumpt_self_past_yn" = "as_is",
                  "wildlife_consumpt_community_past_yn" = "as_is",
                  "wildlife_consumpt_self_curr_cat" = c("tokenize", "combine"),
                  "wildlife_consumpt_self_curr_free" = c("tokenize", "combine"),
                  "wildlife_consumpt_community_curr_cat" = c("tokenize", "combine"),
                  "wildlife_consumpt_community_curr_free" = c("tokenize", "combine"))

cols_meta <- list_of_character_vectors_to_logical_matrix(cols_meta)
colnames(cols_meta)[1] <- "name"

cols_meta
```

## Export columns we're just going to leave as-is

```{r as_is_columns}
as_is <- cols_meta$name[cols_meta$as_is]
cols_clean <- cols_raw[, cols_meta$name[cols_meta$as_is]]

cleaned_cols <- names(cols_clean)

```

## Clean wildlife columns which require tokenizing and converting.

```{r animal_columns}
wildlife_near_home_tax_cat <- cols_raw[, "wildlife_near_home_tax_cat"]
wildlife_in_home_tax_cat <- cols_raw[, "wildlife_in_home_tax_cat"]



col_funs <- function(x) {
  names(x) <- cols_raw$pid

  x <- tokenized_list(x)

  animal_replace = c("rodents" = "rodent",
                     "rats" = "rodent",
                     "shrew-faced ground squirrel" = "rodent",
                     "small mammals" = "small mammal",
                     "foxes" = "small mammal",
                     "bats" = "bat",
                     "birds" = "bird",
                     "primates" = "primate",
                     "reptiles" = "reptile",
                     "snake" = "reptile",
                     "snakes" = "reptile",
                     "other" = "other",
                     "elephants" = "other",
                     "yellow trotted" = "other")

  x <- replace_values(x, animal_replace)

  x <- list_of_character_vectors_to_logical_matrix(x)
}

wildlife_near_home_tax_cat_clean <- col_funs(wildlife_near_home_tax_cat)
wildlife_in_home_tax_cat_clean <- col_funs(wildlife_in_home_tax_cat)

cleaned_cols <- c(cleaned_cols, "wildlife_near_home_tax_cat", "wildlife_in_home_tax_cat")

```

## Clean columns which need just minor replacements.

```{r}
# Look at the names of columns we have yet to clean.
cols_meta2 <- cols_meta[!cols_meta$name %in% cleaned_cols, ]

# These ones:
cols_meta2[1:5, ]

minor_cleaning <- cols_raw[, colnames(cols_raw) %in% cols_meta2$name[1:5]]

for (name in names(minor_cleaning)) {
  cat("\n\n")
  print(name)
  print(descending_kable(minor_cleaning[, name]))
}

names(minor_cleaning)

minor_cleaning[, 1] <- revalue(minor_cleaning[, 1], 
                               replace = c("daily, few times per week" = "other",
                                           "daily, other (specify)" = "other",
                                           "few times per week, other (specify)" = "other"))

minor_cleaning[, 2] <- revalue(minor_cleaning[, 2], 
                               replace = c("bitten, other (specify)" = "bitten",
                                           "bitten, scratched" = "bitten"))

minor_cleaning[, 3] <- revalue(minor_cleaning[, 3], 
                               replace = c("in home, in the forest, other" = "other",
                                           "in the garden" = "in home"))

minor_cleaning[, 4] <- revalue(minor_cleaning[, 4], 
                               replace = c(animal_replace,
                                           "reptile, other" = "reptile",
                                           "leaf snake" = "reptile"))

minor_cleaning[, 5] <- revalue(minor_cleaning[, 5], 
                               replace = c("few times per week, other" = "other",
                                           "few times per month, other" = "other",
                                           "never, daily" = "other"))

for (name in names(minor_cleaning)) {
  cat("\n\n")
  print(name)
  print(descending_kable(minor_cleaning[, name]))
}

cols_clean = cbind(cols_clean, minor_cleaning)

```

## Clean the last four columns, which require tokenizing but not converting.

```{r}
cols_meta3 <- cols_meta[15:18, ]

wildlife_consumpt_self_curr <- cols_raw[, 15:16]
wildlife_consumpt_community_curr <- cols_raw[, 17:18]

col_funs2 <- function(x) {
  x2  <- paste(x[, 1], x[, 2], sep = ",")
  names(x2) <- cols_raw$pid

  x2 <- tokenized_list(x2)

  x2 <- list_of_character_vectors_to_logical_matrix(x2)

  return(x2)
}

wildlife_consumpt_self_curr_clean <- col_funs2(wildlife_consumpt_self_curr)
wildlife_consumpt_community_curr_clean <- col_funs2(wildlife_consumpt_community_curr)

wildlife_consumpt_self_curr_clean
wildlife_consumpt_community_curr_clean

```

## Save all clean CSVs.

```{r}
# Get a list of cleaned data frame names
to_save <- grep("clean$", ls(envir = .GlobalEnv), value = TRUE)

list_df <- mget(grep("clean$", ls(envir = .GlobalEnv), value = TRUE))

lapply(seq_along(list_df),
       function(i) write.csv(list_df[[i]], 
                             paste0("out/", names(list_df)[i], ".csv"),
                             row.names = FALSE))

```